---
title: "Rate of Contraceptive Use in the US"
author: "By: wijube"
date: "2024-02-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About the data

This page was generated with R Language as part of the Digital Health exam in master's programme Global Public Heath. The dataset was taken as an example to learn about Data Visualisation. Target group are students who learn about this topic. Any feedback is welcomed. 

## This is the whole dataset as a Table

```{r message=FALSE, warning=FALSE}
data <- read.csv("CSV-dataset.csv") # Read CSV Data 
knitr::kable(data) # Show the whole data as a table
```
## Filter Table by Year
### List all data for the year 2014
```{r message=FALSE, warning=FALSE}
library(tidyverse) # for ggplot
data <- read.csv("CSV-dataset.csv") # Read CSV Data 
data2014 <- data %>%
    filter(Year == 2014) # Filter the data by year
knitr::kable(data2014) # Show the data of the year 2014
```
### Visualize data for the year 2014 in a bar chart
```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(tidyverse) # for ggplot
data <- read.csv("CSV-dataset.csv") # Read CSV Data 
names(data)[names(data) == "State"] <- "state" # Clean dataset
data2014 <- data %>%
    filter(Year == 2014) # Filter the data by year
# Define color scale
custom_colors <- c("red", "orange", "#49f579")

# Create breaks and labels
breaks <- c(0, 20, 40, max(data2014$RoCU))
labels <- c("0-20", "20-40", "40+")

# Create a bar plot
bar_plot <- ggplot(data2014, aes(x = reorder(state, RoCU), y = RoCU, fill = cut(RoCU, breaks = breaks, labels = labels))) +
  geom_bar(stat = "identity") +
  labs(title = "Rate of Contraceptive Use by State in 2014",
       x = "State",
       y = "Rate of Contraceptive Use",
       fill = "Rate of Contraceptive Use") +
  scale_fill_manual(values = custom_colors) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Print the bar plot
print(bar_plot)
```

### List all data for the year 2015
```{r message=FALSE, warning=FALSE}
library(tidyverse) # for ggplot
data <- read.csv("CSV-dataset.csv") # Read CSV Data 
data2015 <- data %>%
    filter(Year == 2015) # Filter the data by year
knitr::kable(data2015) # Show the data of the year 2015
```
### Visualize data for the year 2015 in a bar chart
```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(tidyverse) # for ggplot
data <- read.csv("CSV-dataset.csv") # Read CSV Data 
names(data)[names(data) == "State"] <- "state" # Clean dataset
data2015 <- data %>%
    filter(Year == 2015) # Filter the data by year
# Define color scale
custom_colors <- c("red", "orange", "#49f579")

# Create breaks and labels
breaks <- c(0, 20, 40, max(data2015$RoCU))
labels <- c("0-20", "20-40", "40+")

# Create a bar plot
bar_plot <- ggplot(data2015, aes(x = reorder(state, RoCU), y = RoCU, fill = cut(RoCU, breaks = breaks, labels = labels))) +
  geom_bar(stat = "identity") +
  labs(title = "Rate of Contraceptive Use by State in 2015",
       x = "State",
       y = "Rate of Contraceptive Use",
       fill = "Rate of Contraceptive Use") +
  scale_fill_manual(values = custom_colors) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Print the bar plot
print(bar_plot)
```

### List all data for the year 2016
```{r message=FALSE, warning=FALSE}
library(tidyverse) # for ggplot
data <- read.csv("CSV-dataset.csv") # Read CSV Data 
data2016 <- data %>%
    filter(Year == 2016) # Filter the data by year
knitr::kable(data2016) # Show the data of the year 2016
```
### Visualize data for the year 2016 in a bar chart
```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(tidyverse) # for ggplot
data <- read.csv("CSV-dataset.csv") # Read CSV Data 
names(data)[names(data) == "State"] <- "state" # Clean dataset
data2016 <- data %>%
    filter(Year == 2016) # Filter the data by year
# Define color scale
custom_colors <- c("red", "orange", "#49f579")

# Create breaks and labels
breaks <- c(0, 20, 40, max(data2016$RoCU))
labels <- c("0-20", "20-40", "40+")

# Create a bar plot
bar_plot <- ggplot(data2016, aes(x = reorder(state, RoCU), y = RoCU, fill = cut(RoCU, breaks = breaks, labels = labels))) +
  geom_bar(stat = "identity") +
  labs(title = "Rate of Contraceptive Use by State in 2016",
       x = "State",
       y = "Rate of Contraceptive Use",
       fill = "Rate of Contraceptive Use") +
  scale_fill_manual(values = custom_colors) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Print the bar plot
print(bar_plot)
```